### 2020-12-08 06_Problems/04_GroupsAgeDistribution/Propensity_score_matching.Rmd
Matching method implementation and regular meeting preparation

---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r, include = F}
library(MatchIt)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(plyr)
library(cobalt)
library(broom)
library(kableExtra)
library(optmatch)
library(rlist)
library(lubridate)
```

```{r}
Sys.setlocale("LC_ALL", "C")
```

```{r}
wd = "/Users/mymac/Documents/research/CFI_Matching/"
setwd(wd)
```

```{r}
df = read.csv(paste0(wd,"/Propensity_score_matchingAllinOne_v2.csv"))
dim(df)
```

```{r}
df$LAD =rep(0, dim(df)[1])
df$LCX =rep(0, dim(df)[1])
df$RCA =rep(0, dim(df)[1])
for (i in 1:dim(df)[1]){
  if (df$Vessel[i] == 0){
    df$LAD[i] = 1
  }
  if (df$Vessel[i] == 1){
    df$LCX[i] = 1
  }
  if (df$Vessel[i] == 2){
    df$RCA[i] = 1
  }
}
df$LAD
df$LCX
df$RCA
```

```{r}
drops <- c("X","X.1", "X.2", "Sex", "Hypertension", "Vessel")
df = df[ , !(names(df) %in% drops)]
colnames(df)
```

```{r}
for (i in 1:dim(df)[1]){
  if (df$Group[i] == 0){
    df$Group[i] = 1
  }
  else {
    df$Group[i] = 0
  }

}
```

```{r, include = F}
df_cov = c(
  "CFI",
  "QCA",
  "Presence.of.AP",
  "LAD",
  "LCX",
  "RCA",
  "Age"
           )
length(df_cov)
```

```{r}
df$Group = as.numeric(df$Group)
```

```{r}
# df$Vessel <- as.factor(df$Vessel)
# df$Vessel
```

```{r, include = F}
sapply(df, class)
```

```{r}
sapply(df, summary)
```

```{r, include = F}
print("-----------------")
for (i in 1:length(df_cov)){
  print(df_cov[i])
  print(table(is.na(df[, df_cov[i]])))
  # for (j in 1:dim(df)[1]){
  #   if(is.na(df[j, df_cov[i]])){
  #     print(df[j, "center"])
  #   }
  #     else{
  #       next
  #     }
  # }
print("-----------------")
  }
```

```{r, include = F}
dim(df)
na_removed_from_df = function(covar) {
  na_var = numeric()
  for (i in 1:length(covar)){
    if ((df[, covar[i]]) %>% anyNA(recursive = FALSE) ==T){
      mess1 = paste0("NA were present in :", covar[i])
      print(mess1)
      na_var = c(na_var, which(is.na(df[,covar[i]]) == T))
    }
    else{
      next
    }
  }
  if(length(na_var) == 0){
    print("there was no NA found")
    df_new = df
  }
  else{
    mes = paste0("entries removed: ")
    mes2 = paste0(na_var, collapse = " ")
    mes_fin = paste0(mes, mes2)
    print(mes_fin)
    df_new = df[-na_var,]
  }
  return(df_new)
}
df_no_na = na_removed_from_df(df_cov)
dim(df_no_na)
```

```{r, include = F}
ttest = t.test(df_no_na$Age ~ df_no_na$Group)
# ttest = t.test(df$age_at_LAAC ~ df$device)
ttest
ttest$estimate[1]
ttest$estimate[2]
```

```{r}
# ttest_table_no_match = sapply(df_cov, function(v){
#   ttest = t.test(df_no_na[, v] ~ df_no_na$device_num)
#   p_value = round(ttest$p.value, digit =3)
#   estimate_asd_pfo = round(ttest$estimate[1], digit = 3)
#   estimate_punct = round(ttest$estimate[2], digit = 3)
#   ttest_table = c(p_value, estimate_asd_pfo, estimate_punct)
#   names(ttest_table) = c("p value", "estimate_asd_pfo", "estimate_punct")
#   return(ttest_table)
#   })
# ttest_table_no_match = t(as.data.frame(ttest_table_no_match))
# ttest_table_no_match = ttest_table_no_match[, c(2,3,1)]
# ttest_table_no_match
```

```{r, include = F}
# tab1 = kable(ttest_table_no_match, "latex", booktabs = T, col.names = c("estimate", "st. error", "p-value")) %>% 
# add_header_above(c("Unpaired t-test on variables on\namplatzer vs lambre - not matched" = 4)) %>%
# kable_styling(latex_options = c("striped", "hold_position"))
# kable_as_image(tab1, "/Users/mymac/Documents/research/Lambre_vs_Amplatzer/figures_and_tables/0_ttest_not_matched", file_format = "png")
```

```{r, include = F}
ps_df_not_matched = glm(f.build("Group", df_cov),
                      family = binomial(), data = df_no_na)
summa_ps_df_not_matched = summary(ps_df_not_matched)
df_not_matched_logit_coef = summa_ps_df_not_matched$coefficients
df_not_matched_logit_coef = round(df_not_matched_logit_coef, digit = 4)
df_not_matched_logit_coef = subset(df_not_matched_logit_coef, select = -c(3))
df_not_matched_logit_coef
```

```{r, include = F}
tab2 = kable(df_not_matched_logit_coef, "latex", booktabs = T, col.names = c("estimate", "st. error", "p-value")) %>% 
add_header_above(c("Logistic regression of variables on\namplatzer vs lambre - not matched" = 4)) %>%
# column_spec(1, width = "3cm") %>%
kable_styling(latex_options = c("striped", "hold_position"))
# kable_as_image(tab2, "/Users/mymac/Documents/research/Lambre_vs_Amplatzer/figures_and_tables/0_log_reg_not_matched", file_format = "png")
```

```{r, include = F}
prs_df = data.frame(pr_score = predict(ps_df_not_matched, type = "response"),
                    device = ps_df_not_matched$model$Group)
```

```{r, include = F}
labs = paste("actual procedure:", c("lambre", "amplatzer"))
prs_df %>%
  mutate(device = ifelse(device == 1, labs[1], labs[2])) %>%
  ggplot(aes(x = pr_score)) +
  geom_histogram(color = "white") +
  facet_wrap( ~ device) +
  xlab("Probability of Puncture") +
  theme_bw()
```

```{r}
df_mod_selec = data.frame(matrix(ncol = 4, nrow = 0))
col_names = c("model", "mean diff", "ratio", "caliper")
colnames(df_mod_selec) = col_names
df_mod_selec
```
```{r}
colnames(df_no_na)
```

```{r}
df_cov
```

```{r, echo = F}

#df_mod_selec = data.frame(matrix(ncol = 4, nrow = 0))
#col_names = c("model", "mean diff", "ratio", "caliper")
#colnames(df_mod_selec) = col_names
#df_mod_selec

col_names = c("model", "mean diff", "ratio", "caliper", "n_control_matched", "n_patients")
df_mod_selec = data.frame(matrix(ncol = length(col_names), nrow = 0))
colnames(df_mod_selec) = col_names
iter = 500
threshold = 0.60
ratio_list = c(1, 2, 3)
caliper_list = c(0.8, 0.6, 0.4, 0.2)
#ratio_list = c(2)
#caliper_list = c(0.4)
for (j in 1:length(ratio_list)){
  ratio_val = ratio_list[j]
  print(paste0("ratio: ", ratio_val))
  for (i in 1:length(caliper_list)){
      caliper_val = caliper_list[i]
      print(paste0("caliper: ", caliper_val))

    select_mod = data.frame(matrix(ncol = length(df_cov) + 2, nrow = iter))
    colnames(select_mod) = c(df_cov, "mean", "seed")
    dim(select_mod)
    for (i in 1:iter){
      set.seed(i)
      mod_match_logit = matchit(f.build("Group", df_cov), method = "nearest", distance = "logit", data = df_no_na, caliper = caliper_val, ratio = ratio_val)
      tabl = bal.tab(mod_match_logit)
      df_tbl_match_logit = data.frame(tabl$Balance$Diff.Adj)
      rownames(df_tbl_match_logit) = rownames(tabl$Balance)
      df_tbl_match_logit = rbind(df_tbl_match_logit, "Mean" = mean(abs(tabl$Balance$Diff.Adj)))
      df_tbl_match_logit = rbind(df_tbl_match_logit, "seed" = i)
      for (j in 2:dim(df_tbl_match_logit)[1]){
          select_mod[i,j-1] = df_tbl_match_logit[j,1]
      }
    }
    # create a new df without variable "mean" and "seed" for the function all()
    select_mod_red = select_mod[, -which(names(select_mod) %in% c("mean", "seed"))]

    # select model with all standardized difference < threshold
    select_index = c()
    for (i in 1:dim(select_mod_red)[1]){
      if(all(select_mod_red[i,] < threshold) == T){
        select_index = c(select_index, i)
      }
      else{
        next
      }
    }

    select_mod_ind = select_mod[select_index, ]

    golden_seed = select_mod_ind$seed[which(select_mod_ind$mean == min(select_mod_ind$mean))]
    set.seed(golden_seed)
    mod_match_logit = matchit(f.build("Group", df_cov), method = "nearest", distance = "logit", data = df_no_na, caliper = caliper_val, ratio = ratio_val, replace = FALSE)
    tabl = bal.tab(mod_match_logit)

    mean_diff = mean(abs(tabl$Balance$Diff.Adj))

    model_str = paste0("ratio: ", ratio_val, " - caliper ", caliper_val, ":1")

     df_mod_added = data.frame(model_str, mean_diff, ratio_val, caliper_val,  tabl$Observations$Treated[2], tabl$Observations$Control[2])
    #df_mod_added = data.frame(model_str, mean_diff, ratio_val, caliper_val)
    df_mod_selec = rbind(df_mod_selec, df_mod_added)
    df_mod_selec

    #df_observation = data.frame(rownames(tabl$Balance), tabl$Balance[,4])
    #colnames(df_balance) = c("variable", "Standardized difference")
    write.csv(tabl$Observations, paste0(wd, "Observations/observations_ratio_", ratio_val, "_caliper_", caliper_val,  ".csv"))

    df_balance = data.frame(rownames(tabl$Balance), tabl$Balance[,4])
    colnames(df_balance) = c("variable", "Standardized difference")
    write.csv(df_balance, paste0(wd, "balance/balance_ratio_", ratio_val, "_caliper_", caliper_val,  ".csv"), row.names = FALSE)

    love1 = love.plot(bal.tab(mod_match_logit), threshold = .1)
    ggsave(love1, file = paste0("love_ratio_", ratio_val, "_caliper_", caliper_val, ".png"), path = paste0(wd, "/figures_and_tables"))

    df_m_logit = match.data(mod_match_logit)
    for (i in 1:dim(df_m_logit)[1]){
      if (df_m_logit$Group[i] == 0){
        df_m_logit$Group[i] = 1
      }
      else {
        df_m_logit$Group[i] = 0
      }
    }

    ps_df_match_logit = glm(f.build("Group", df_cov), family = binomial(), data = df_m_logit)
    summa_ps_df_match_logit = summary(ps_df_match_logit)
    df_m_logit_coef = summa_ps_df_match_logit$coefficients
    df_m_logit_coef = round(df_m_logit_coef, digit = 4)
    df_m_logit_coef = subset(df_m_logit_coef, select = -c(3))
    df_m_logit_coef
    write.csv(df_m_logit_coef, paste0(wd, "lr_baseline_characteristcs_post_matching/logistic_regression_ratio_", ratio_val, "_caliper_", caliper_val, ".csv"))

    df_m_logit$Vessel =rep(0, dim(df_m_logit)[1])
    for (i in 1:dim(df_m_logit)[1]){
      if (df_m_logit$LAD[i] == 1){
        df_m_logit$Vessel[i] = 0
      }
      if (df_m_logit$LCX[i] == 1){
        df_m_logit$Vessel[i] = 1
      }
      if (df_m_logit$RCA[i] == 1){
        df_m_logit$Vessel[i] = 2
      }
    }
    df_m_logit$LAD <- NULL
    df_m_logit$LCX <- NULL
    df_m_logit$RCA <- NULL
    #df_m_logit$distance <- NULL
    df_m_logit$weights <- NULL

    write.csv(df_m_logit, paste0(wd, "matched_data/matched_data_ratio_", ratio_val, "_caliper_", caliper_val, ".csv"))
  }
}
colnames(df_mod_selec) = col_names
write.csv(df_mod_selec, paste0(wd, "/df_mod_selec.csv"))
```